---
title: "Absolute Cell Counting"
output: html_notebook
---

```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
options(warn = -1)
root <- rprojroot::find_root(".git/index");  source(file.path(root, "src/init.R"))
shawn_data_files <- "/home/data/refined/reef/microscopy/re_analysis/"
analysis_files <- "/home/data/refined/reef/microscopy/analysis_Nov_28_2020"
library(tidyverse)
library(prodlim)
```

```{r, warning=FALSE,message=FALSE,error=FALSE}
raw_images <- "/home/data/refined/reef/microscopy/imagej_analysis/re-analysis"
batch_dirs <-         c( "20200901_bellairs", "20200901_maycocks")
master <- tibble()
```



```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
target <- 1
fls <- list.files(file.path(raw_images, batch_dirs[target]),
                  pattern = "*.csv")

data <- lapply( 1:length(fls), FUN = function(x) {
  tmp <- read_csv(file.path( raw_images, batch_dirs[target], fls[x] ))
  tmp$site <- "B"
  tmp$file <- x
  return(tmp)
})
data <- bind_rows(data)
master <- bind_rows(master, data)
```



```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
target <- 2
fls <- list.files(file.path(raw_images, batch_dirs[target]),
                  pattern = "*.csv")

data <- lapply( 1:length(fls), FUN = function(x) {
  tmp <- read_csv(file.path( raw_images, batch_dirs[target], fls[x] ))
  tmp$site <- "M"
  tmp$file <- x
  return(tmp)
})
data <- bind_rows(data)
master <- bind_rows(master, data)
```

Reorganize slightly and save.
```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
master <- master %>%  mutate( FOV = X1, .keep="unused") %>% dplyr::select(site, file, FOV, everything())
#saveRDS(master, file = file.path(analysis_files, "object_tibble_vers_1.0.RData"))
```


```{r}
master <- readRDS(file.path(analysis_files, "object_tibble_vers_1.0.RData") )
p <- ggplot(master, aes(factor(site), log(Area)))
p + geom_violin()
```

Not soo informtive yet. Let's look at specific size regions.

1) Prochlorocous size : (0.6 to 0.8 μm in diameter) - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6391267/

2) Synechococcus size : (0.9–1.7 µm in diameter) - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6391267/

3) Pelagibacteria size : (0.2 μm in diameter) - link above in previous size labelling

```{r}
target_areas <- c(  pi*(0.2/2)^2, pi*(0.8/2)^2, pi * (1.7/2)^2, pi * (3/2)^2 ) 
master %>% filter(site=="B", Area < target_areas[4]) -> B_target; summary(B_target$Area)
master %>% filter(site=="M", Area < target_areas[4]) -> M_target; summary(M_target$Area)
```
```{r}
target <- master %>% filter( Area < target_areas[4] )
p <- ggplot(target, aes(factor(site), Area))
p + geom_violin()
```


```{r}
target <- master %>% filter(  Area > target_areas[2], Area < target_areas[3])
p <- ggplot(target, aes(factor(site), Area))
p + geom_violin()
```


Let's look at just Pelagibachter

```{r}
target_areas_pelag <- c( pi * (0.2/2)^2, pi * (0.5/2)^2 )
target <- master %>% filter(  (Area > target_areas_pelag[1] & Area < target_areas_pelag[2]) )
p <- ggplot(target, aes(factor(site), Area))
p + geom_violin()
```


Let's look at just Synech.

```{r}
target_areas_synech <- c( pi * (0.9/2)^2, pi * (1.8/2)^2 )
target <- master %>% filter(  (Area > target_areas_synech[1] & Area < target_areas_synech[2]) )
p <- ggplot(target, aes(factor(site), Area))
p + geom_violin()
```


Let's look at just Proch.

```{r}
target_areas_proch <- c( pi * (0.5/2)^2, pi * (0.9/2)^2)
target <- master %>% filter(  (Area > target_areas_proch[1] & Area < target_areas_proch[2]) )
p <- ggplot(target, aes(factor(site), Area))
p + geom_violin()
```




```{r}
target_areas <- c(  pi*(0.2/2)^2, pi*(0.8/2)^2, pi * (1.7/2)^2, pi * (3/2)^2 ) 
target <- master %>% filter( Area > target_areas[1], Area < target_areas[4])
ggplot( target, aes(x=log(Area), color = site)) + 
  geom_histogram( aes(y=..density..), 
    fill="white", alpha=0.5, position="identity", bins=100) +
  geom_density(alpha=0.6)+
  theme_classic() +
  theme(legend.position="top")
```

```{r}
target_areas <- c(  pi*(0.2/2)^2, pi*(0.8/2)^2, pi * (1.7/2)^2, pi * (3/2)^2 ) 

target_b <- master %>% filter( Area > target_areas[1], Area < target_areas[4])
bb <- ggplot( target_b, aes(x=log(Area), color = site)) + 
  geom_histogram( aes(y=..density..), 
    fill="white", alpha=0.5, position="identity", bins=100) +
  geom_density(alpha=0.6)+
  theme_classic() +
  theme(legend.position="top")
bb
```

```{r}
summary(target_b3$Area)
```

Get the basic counts that we need.

```{r}
(res <- master %>% filter( Area > target_areas[1], Area < target_areas[4]) %>%
  group_by( site,  file ) %>%
    summarise( n= n(),  mean_area = mean(Area), StdDev = sd(Area, na.rm=TRUE)  ) )
```




```{r}

ggplot(res) + geom_point( aes( x= mean_area, y= StdDev, color =  site ))
```


I think we should thow out the outlier here (n) but first let's investigate the most extreme a little bit more.
```{r}
big <- res %>% filter(n > 3000) %>% dplyr::select(site, batch, file) # let's pick on big[2,]

reduced <- master
for (i in 1:nrow(big)) {
  reduced <- reduced %>% filter( !( site == as.character(big[i,1]) &
                                    batch == as.numeric(big[i,2]) & 
                                    file == as.numeric(big[i,3])))
}

reduced <- reduced  %>% 
    filter( Area > target_areas[1]) %>%
    filter( Area < target_areas[4]) 

ggplot(reduced, aes(x=log(Area), color = site)) + 
  geom_histogram( aes(y=..density..), 
    fill="white", alpha=0.5, position="identity", bins=100) +
  geom_density(alpha=0.6)+
  theme_classic() +
  theme(legend.position="top")
```

What is the average number of FOV's per file at each site?

```{r}
res <- reduced %>% group_by(site, batch, file) %>%
    summarise( n= n(), mean_area= mean(Area))
res2 <- res %>% group_by(site) %>%
    summarise(nn = n(), ave_n = mean(n))

B_n <- res %>% filter(site == "B")
M_n <- res %>% filter(site == "M")
t.test(B_n$n, M_n$n)
```

```{r}

ggplot(res, aes(x=n, color = site)) + 
  geom_histogram( aes(y=..density..), 
    fill="white", alpha=0.5, position="identity", bins=100) +
  geom_density(alpha=0.6)+
  theme_classic() +
  theme(legend.position="top")
```
